### 2024-03-16 Response/Response to reviewer.tex
Praveen-16-March-24
Response to reviewer-By Praveen

\documentclass[12pt,twoside,reqno]{amsart}
%\documentclass[12pt]{article}

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{amsthm, amsmath, amscd, amssymb,centernot}
\usepackage[all]{xy}
\usepackage[T1]{fontenc}
\usepackage[left=2.5cm,top=2.5cm,bottom=3cm,right=2.5cm]{geometry}
\usepackage{amsthm, amsmath, amscd, amssymb,centernot}
\usepackage{comment}
\setlength{\headheight}{15.2pt}
\setlength\parskip{.1in}
\setlength\parindent{0.2in}

%\usepackage{tikz}https://www.overleaf.com/project/642e283246ec4730b8bdf398
\usepackage{csquotes}

\newcommand{\s}{\varepsilon}

\numberwithin{equation}{section}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{definition}
\newtheorem{question}[theorem]{Question}

\begin{document}
\title{Response to Reviewer}
\begin{description}
\item [Mauscript Number] GMJ 231005 \\
\item [Full Title] Seshadri Constants of Curve Configurations on Surfaces
\item [Author] Krishna Hanumanthu, Praveen Roy and Aditya Subramaniam
\end{description}

We start by thanking the anonymous referee for carefully reading the manuscript and suggesting 
useful changes and also for pointing out an error in the Assumption 2.1 and therefore the ambiguity in Theorem 2.5. We also agree with the reviewer that the example 2.21 is 
better to be shorten focusing only on the computation related to Seshadri constants. We have made all these changes, and we hope that after making these changes, the paper has becomes more accessible for the readers and its overall quality has improved considerably.  Following are the detailed (point by point) explanation of changes that we have made on the basis of the report. \\

We agree with the referee about the problem in our way of using the Assumption 2.1. While writing the paper, we always had in mind the \textbf{connectedness of the arrangement}, 
which we somehow missed to incorporate in the Assumption 2.1. We have made it clear in the manuscript now. Note that, now we don't require the semi-ampleness of divisor $A$ for $C_i^2$ to be positive, and it follows 
from the connectedness of the arrangement. Also, we would like to keep second assumption in Assumption 2.1, as removing that requires lot of new work and we aren't quite interested to carry out the required work, in that direction, 
as of now. The last concern of reviewer has also been addressed, we 
have updated the last example 2.21 and it has been shorten. \\

We once again thank the referee for a quick review, we hope that we addressed all the concern of the referee regarding this manuscript.

\end{document}